A linear inside-outside algorithm for correcting sequencing errors in structured RNA sequences

### 2012-10-07 Recomb/introduction_RECOMB.tex
fix ref

For half a century, biological molecules have been studied as a proxy to understand evolution~\cite{Zuckerkandl1965}, and due
to their fundamental functions and remarkably conserved structures, rRNAs have always been a prime candidate for phylogenetic
studies~\cite{Olsen1986, Olsen1993}. In recent years, studies as the \emph{Human Microbiome Project}~\cite{Turnbaugh2007} benefited
of new technologies such as the NGS techniques to sequence as many new organisms as possible and extract a unprecedented flow of new information. 
Nonetheless, these high-throughput techniques typically have high error rates that make their applications to metagenomics (a.k.a. environmental
genomics) studies challenging. For instance, pyrosequencing as implemented by Roche's 454 produces may have a error rate raising up to 10\%.
Because there is no cloning step, resequencing to increase accuracy is not possible and it is therefore vital to disentangle noise from true sequence
diversity in this type of data \cite{Quince:2009uq}. Errors can be significantly reduced  when large multiple sequence alignments with close homologs
are available, but in studies of new or not well known organisms, such information is rather sparse. In particular, it is common to is not enough  similarity to 
differentiate between the sequencing errors and the natural polymorphisms that we want to observe, often inflating the diversity estimates~\cite{Kunin2010}.
A few techniques have been developed to remedy to this problem~\cite{Quinlan2010,Medvedev2011} but they do not take
into account all the available information. It is therefore essential to develop methods that can exploit any type of signal available to correct errors.

>>>>

For half a century, biological molecules have been studied as a proxy to understand evolution~\cite{Zuckerkandl1965}, and due
to their fundamental functions and remarkably conserved structures, rRNAs have always been a prime candidate for phylogenetic
studies~\cite{Olsen1986, Olsen1993}. In recent years, studies as the \emph{Human Microbiome Project}~\cite{Turnbaugh2007} benefited
of new technologies such as the NGS techniques to sequence as many new organisms as possible and extract a unprecedented flow of new information. 
Nonetheless, these high-throughput techniques typically have high error rates that make their applications to metagenomics (a.k.a. environmental
genomics) studies challenging. For instance, pyrosequencing as implemented by Roche's 454 produces may have a error rate raising up to 10\%.
Because there is no cloning step, resequencing to increase accuracy is not possible and it is therefore vital to disentangle noise from true sequence
diversity in this type of data \cite{Quince:2009uq}. Errors can be significantly reduced  when large multiple sequence alignments with close homologs
are available, but in studies of new or not well known organisms, such information is rather sparse. In particular, it is common to is not enough  similarity to 
differentiate between the sequencing errors and the natural polymorphisms that we want to observe, often inflating the diversity estimates~\cite{Kunin2010}.
A few techniques have been developed to remedy to this problem~\cite{Quinlan2008,Medvedev2011} but they do not take
into account all the available information. It is therefore essential to develop methods that can exploit any type of signal available to correct errors.
